---
title: "Crime-Space"
author: "Zheyan"
date: "11/9/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide 
    theme: cosmo

---

```{r setup, include=FALSE}
library(leaflet)
library(tidyverse)
library(plotly)
##install.packages("leaflet.extras")
library(leaflet.extras)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


```{r read,include=FALSE}

crime_df = read_csv('subwaycrime_with_station_new_2.csv')

crime_df<-crime_df %>% 
 mutate(cmplnt_to_dt=as.Date(cmplnt_to_dt,format='%m/%d/%Y')) %>% 
 filter(distance<=0.0001)%>% 
 filter(cmplnt_to_dt>='2006-1-1')

```

## Where crime?

New York City can be a dangerous place and crime from above ground will 
often extend into the NYC Subway.</br>
We mainly focus on the recent crime data on subway in NYC in this year, and 
there are `r nrow(crime_df)` complaints from 2006 to now. 

## Heat Map of Subway Crime in NYC, 2006-2021

From this map, you can check where the crime happened frequently. 
```{r heatmap, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  group_by(closest_station) %>% 
  summarise(Trend=n(),
            longitude=mean(longitude),
            latitude=mean(latitude)) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addHeatmap(lng=~longitude,lat=~latitude,
             intensity=~exp(Trend),max=100,radius=5,blur=10)

```

## Map of Subway Crime in NYC, 2006-2021

From this map, you can check each crime's location, type, victim and suspects' 
information and time. 

```{r leaflet, echo = FALSE, message=FALSE, warning=FALSE}



leaflet(crime_df) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~latitude, lng = ~longitude, 
             popup = paste("Offense Description:", crime_df$ofns_desc, "<br>",
                           "Suspect's age:",crime_df$susp_age_group,"<br>",
                           "Suspect's race:", crime_df$susp_race,"<br>",
                           "Suspect's gender:",crime_df$susp_sex,"<br>",
                           "Victim's age:",crime_df$vic_age_group,"<br>",
                           "Victim's race:", crime_df$vic_race,"<br>",
                           "Victim's gender:",crime_df$vic_sex,"<br>",
                           "Time:",crime_df$cmplnt_to_dt),
                clusterOptions = markerClusterOptions())
  


# "Offense Description:", crime_df$ofns_desc, "<br>",
# "<br>", 
#                 
#                 "Suspect Information", c(crime_df$susp_age_group,
#                                          crime_df$susp_race,
#                                          crime_df$susp_sex), "<br>",
#                 "Victim Information:", c(crime_df$vic_age_group,
#                                          crime_df$vic_race,
#                                          crime_df$vic_sex), 
```

## Distribution of crime in 7 Clusters

When firstly scanning the map above, you can ambiguously know how many crimes 
in each part of NYC, so let us check them in each ambiguously using bar chart.

```{r barcluster, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  count(cluster) %>% 
  mutate(cluster = fct_reorder(cluster, n)) %>% 
  slice_max(n,n=20) %>% 
  plot_ly(x = ~cluster, y = ~n, color = ~cluster, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Cluster'))
```

## Top 10 offense classification


There are `r count(distinct(crime_df,ofns_desc))` kinds of crime occurring in the subway, there is the bar chart shows the wildest 10 crimes in the subway.

The most frequent crime mainly consists of grand larceny and assaults.

```{r offense, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  count(ofns_desc) %>% 
  mutate(ofns_desc = fct_reorder(ofns_desc, n)) %>% 
  slice_max(n,n=10) %>% 
  plot_ly(x = ~ofns_desc, y = ~n, color = ~ofns_desc, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Type of Offense'))

## some data's resources
## https://stackoverflow.com/questions/51955803/how-to-change-x-axis-layout-using-plotly-in-r

```

## Top 20 station where crime happens frequently

From this chart, you can mainly check which station is the most dangerous station.
```{r station, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  count(closest_station) %>% 
  mutate(closest_station = fct_reorder(closest_station, n)) %>% 
  slice_max(n,n=20) %>% 
  plot_ly(x = ~closest_station, y = ~n, color = ~closest_station, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Type of station'))


```




```{r suprace, echo = FALSE, message=FALSE, warning=FALSE}
# 
# crime_df %>% 
#   drop_na(cluster,susp_race) %>% 
#   filter(susp_race!="UNKNOWN") %>% 
#   count(cluster,susp_race) %>% 
#   mutate(cluster = fct_reorder(cluster, n)) %>% 
#   plot_ly(x = ~cluster, y = ~n, color = ~susp_race, type = "bar", colors = "viridis") %>% 
#   layout(
#          yaxis = list(title = 'Number of Compliants'),
#          xaxis = list(title = 'Area'))

```

## Barchart by each Borough about Victims

From this graph, you can check in each borough, which races more possibly 
vulnerable in the subway. 

As you can see, the proportion of African Americans in each cluster are stable; In Manhattan, where has the most crimes, white people(including white hispanic) are more vunlerable than African Americans in these places. 
```{r vicrace, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>%
  filter(vic_race!="UNKNOWN") %>%
  count(cluster,vic_race) %>%
  mutate(cluster = fct_reorder(cluster, n)) %>% 
  plot_ly(x = ~cluster, y = ~n, color = ~vic_race, type = "bar", colors = "viridis") %>%
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Cluster'))

```


## Gender Distrubution for popular crime types

In this part, we want to know which gender are more possible to become potential victims. We choose several types of crime from the most frequent kinds of crimes:</br>
ASSAULT 3 & RELATED OFFENSES,HARRASSMENT 2,GRAND LARCENY,FELONY ASSAULT,
ROBBERY,PETIT LARCENY,SEX CRIMES.</br>
We can find that crimes about sex crimes and harassment, female are more possible to be attacked. Crimes about regular theft, such as assault and larceny, male are potential victims comparing with female. 

```{r gender, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(ofns_desc %in% c("ASSAULT 3 & RELATED OFFENSES",
                          "HARRASSMENT 2","GRAND LARCENY",
                          "FELONY ASSAULT",
                          "ROBBERY","PETIT LARCENY",
                          "SEX CRIMES")
  ) %>% 
  filter(vic_sex %in% c("F","M")) %>% 
  count(ofns_desc,vic_sex) %>% 
  mutate(ofns_desc = fct_reorder(ofns_desc, n)) %>% 
  plot_ly(x = ~ofns_desc, y = ~n, color = ~vic_sex, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Age Group'))
  
                

```


## Female Age Distribution for Sex Crimes 

Let us talk more about age group for some specific crimes: SEX CRIMES and HARRASSMENT 2. </br>
Most of the victims' age are in the age 25-44 interval. 



```{r harrass, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(ofns_desc %in% c("SEX CRIMES","HARRASSMENT 2")) %>% 
  count(vic_age_group,vic_race) %>% 
  filter(vic_age_group!="-59") %>% 
  filter(vic_age_group!="UNKNOWN") %>%
  filter(vic_race!="UNKNOWN") %>%
  plot_ly(x = ~vic_age_group, y = ~n, color = ~vic_race, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Age Group-Female'))
  
```

## Crimr Rate Top 20
Sometimes, we more care about the crime rates on subway rather than number of crimes, because we also care the possiblity that the people standing in front of us is suspective. 
```{r crime_rate, echo = FALSE, message=FALSE, warning=FALSE}
crime_rate_summary<-read.csv("data/crime_rate_station.csv")
crime_rate_summary %>% 
  mutate(crime=n) %>% 
  select(-n) %>% 
  relocate(linename,station,flow,crime) %>% 
  slice_max(crime_rate,n=20) %>%
  knitr::kable()
  
```
