---
title: "EDA"
author: "Zheyan"
date: "11/9/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide 
    theme: cosmo

---

```{r setup, include=FALSE}
library(leaflet)
library(tidyverse)
library(plotly)
##install.packages("leaflet.extras")
library(leaflet.extras)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


```{r,include=FALSE}
# df = read_csv('data/subwaycrime.csv') %>% 
#   janitor::clean_names() %>% 
# df %>% 
#  mutate(cmplnt_to_dt=as.Date(cmplnt_to_dt,format="％M/%D/%Y")) %>% 
#   
#   
# 
# df %>% 
#   leaflet() %>% 
#   addTiles() %>% 
#   addCircleMarkers(~longitude, ~latitude,radius= 3, weight= 0.5,
#                    fillColor=~(boro_nm))

crime_df = read_csv('data/subwaycrime_with_station_new_2.csv')

crime_df<-crime_df %>% 
 mutate(cmplnt_to_dt=as.Date(cmplnt_to_dt,format='%m/%d/%Y')) %>% 
 filter(cmplnt_to_dt>='2021-1-1')

```

## Where crime?

New York City can be a dangerous place and crime from above ground will 
often extend into the NYC Subway.</br>
We mainly focus on the recent crime data on subway in NYC in this year, and 
there are `r nrow(crime_df)` complaints in 2021. 

## Heat Map of Subway Crime in NYC, 2021

From this map, you can check where the crime happened frequently. 
```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  group_by(closest_station) %>% 
  summarise(Trend=n(),
            longitude=mean(longitude),
            latitude=mean(latitude)) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addHeatmap(lng=~longitude,lat=~latitude,
             intensity=~exp(Trend),max=100,radius=5,blur=10)

```

## Map of Subway Crime in NYC, 2021

From this map, you can check each crime's location, type, victim and suspects' 
information and time. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}



leaflet(crime_df) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~latitude, lng = ~longitude, 
             popup = paste("Offense Description:", crime_df$ofns_desc, "<br>",
                           "Suspect's age:",crime_df$susp_age_group,"<br>",
                           "Suspect's race:", crime_df$susp_race,"<br>",
                           "Suspect's gender:",crime_df$susp_sex,"<br>",
                           "Victim's age:",crime_df$vic_age_group,"<br>",
                           "Victim's race:", crime_df$vic_race,"<br>",
                           "Victim's gender:",crime_df$vic_sex,"<br>",
                           "Time:",crime_df$cmplnt_to_dt),
                clusterOptions = markerClusterOptions())
  


# "Offense Description:", crime_df$ofns_desc, "<br>",
# "<br>", 
#                 
#                 "Suspect Information", c(crime_df$susp_age_group,
#                                          crime_df$susp_race,
#                                          crime_df$susp_sex), "<br>",
#                 "Victim Information:", c(crime_df$vic_age_group,
#                                          crime_df$vic_race,
#                                          crime_df$vic_sex), 
```

## Distribution of crime in boroughs

When firstly scanning the map above, you can ambigouously know how many crimes 
in each part of NYc, so let uscheck them in each borough using bar chart.

```{r, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  count(boro_nm) %>% 
  mutate(boro_nm = fct_reorder(boro_nm, n)) %>% 
  slice_max(n,n=20) %>% 
  plot_ly(x = ~boro_nm, y = ~n, color = ~boro_nm, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Area'))
```

## Top 10 offense classification

```{r, echo = FALSE, message=FALSE, warning=FALSE}
## crime_df = read_csv('data/subwaycrime_with_station.csv')

```

There are `r count(distinct(crime_df,ofns_desc))` kinds of crime occurring in the subway, there is the bar chart shows the wildest 10 crimes in the subway.

The most frequent crime mainly consists of sexual assaults and thefts. The assualt
has 869 cases, which is more than a quarter of the total crimes. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  count(ofns_desc) %>% 
  mutate(ofns_desc = fct_reorder(ofns_desc, n)) %>% 
  slice_max(n,n=10) %>% 
  plot_ly(x = ~ofns_desc, y = ~n, color = ~ofns_desc, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Type of Offense'))

## some data's resources
## https://stackoverflow.com/questions/51955803/how-to-change-x-axis-layout-using-plotly-in-r

```

## Top 20 station where crime happens frequently

From this chart, you can mainly check which station is the most dangerous station
in 2021. 
```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df = read_csv('data/subwaycrime_with_station.csv')
crime_df %>% 
  count(closest_station) %>% 
  mutate(closest_station = fct_reorder(closest_station, n)) %>% 
  slice_max(n,n=20) %>% 
  plot_ly(x = ~closest_station, y = ~n, color = ~closest_station, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Type of station'))


```



## Barchart by each Borough about suspects

From this graph, you can check in each borough, which races more possibly 
commit a crime in the subway. 
```{r, echo = FALSE, message=FALSE, warning=FALSE}

crime_df %>% 
  drop_na(boro_nm,susp_race) %>% 
  count(boro_nm,susp_race) %>% 
  plot_ly(x = ~boro_nm, y = ~n, color = ~susp_race, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Area'))

```

## Barchart by each Borough about Victims

From this graph, you can check in each borough, which races more possibly 
vulnerable in the subway. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(vic_race!="UNKNOWN") %>% 
  count(boro_nm,vic_race) %>% 
  plot_ly(x = ~boro_nm, y = ~n, color = ~vic_race, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Area'))

```

## 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(ofns_desc %in% c("SEX CRIMES","HARRASSMENT 2","ASSAULT 3 & RELATED OFFENSES")) %>% 
  count(vic_age_group,vic_race) %>% 
  plot_ly(x = ~vic_age_group, y = ~n, color = ~vic_race, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Age Group'))
  
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(ofns_desc %in% c("ASSAULT 3 & RELATED OFFENSES")) %>% 
  count(vic_age_group,vic_sex) %>% 
  plot_ly(x = ~vic_age_group, y = ~n, color = ~vic_sex, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Age Group'))
  

```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
crime_df %>% 
  filter(ofns_desc %in% c("ASSAULT 3 & RELATED OFFENSES",
                          "HARRASSMENT 2","GRAND LARCENY",
                          "DANGEROUS DRUGS","FELONY ASSAULT",
                          "ROBBERY","PETIT LARCENY","FORGERY",
                          "SEX CRIMES")
  ) %>% count(ofns_desc,vic_sex) %>% 
  
    plot_ly(x = ~ofns_desc, y = ~n, color = ~vic_sex, type = "bar", colors = "viridis") %>% 
  layout(
         yaxis = list(title = 'Number of Compliants'),
         xaxis = list(title = 'Age Group'))
  
                

```
