---
title: "time data"
output: 
  html_document:  
    toc: true 
    toc_float: true 
    code_folding: hide 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(plotly)
library(gganimate)
library(animation)
library(lubridate)
library(leaflet)
ani.options(interval=2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Data Cleanning

### Passenger Data
```{r echo=FALSE,tidy=TRUE,message=FALSE}
# read passgenger data, create unique variable for each station(station_line),get the start and end time of each oberservation
passenger_df <- 
  read_csv("data/passenger_imputed.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    station_line = str_c(station,linename,sep = " line:"),
    start_time = time-14400,
  ) %>% 
  rename(
    end_time = time
  ) %>% 
  select(station_line,station,linename,date,time_period,start_time,end_time,entry_diff_imputed,exit_diff_imputed)

# get the information of each station
subway_info <- 
  read_csv("data/subway_info_final3.csv") %>% 
  mutate(
    station_line = str_c(station,linename,sep = " line:")
    ) %>% 
  select(station_line,service)

# set the color of each station by its line
linecolor <- colorFactor(
  palette = c("blue", "azure4", "orange",'green','brown','yellow','red','forestgreen','purple'),
  levels = c('8 Avenue(ACE)',
             'Shuttle(S)',
             '6 Avenue(BDFM)',
             'Brooklyn-Queens Crosstown(G)',
             '14 St-Canarsie(L)',
             'Broadway(NQRW)',
             '7 Avenue(123)',
             'Lexington Av(456)',
             'Flushing(7)'))


            
```



```{r eval=FALSE,echo=FALSE,tidy=TRUE,message=FALSE}
{start_date <-  as.Date(readline(prompt = "Enter The Start Date(YYYY-MM-DD):"));
end_date <-  as.Date(readline(prompt = "Enter The End Date(YYYY-MM-DD):"));
start_time <- readline(prompt = "Enter the Start Time(00:00,04:00,08:00,12:00,16:00,20:00):");
end_time <- readline(prompt = "Enter the Start Time(00:00,04:00,08:00,12:00,16:00,20:00):")}
```

### Time 
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
# 2021-01-01
start_date <-  ymd("2021-01-01")
end_date <-  ymd("2021-06-30")
start_time <- "00:00"
end_time <- "20:00"

time_s <- as.numeric(substr(start_time,1,2))*3600
time_e <- as.numeric(substr(end_time,1,2))*3600
date_series <- seq(start_date,end_date,by=1)

passenger_df_time <- 
  passenger_df%>% 
  filter(
    date %in% date_series,
    !(date == start_date&start_time<time_s),
    !(date == end_date&start_time>time_e)
  )
```

# Busiest Station 

## Top 10 busiest station
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
passenger_df_time %>% 
  group_by(station_line) %>% 
  summarize(
    total_entry = sum(entry_diff_imputed,na.rm = TRUE),
    total_exit = sum(exit_diff_imputed,na.rm = TRUE),
  ) %>% 
  mutate(total = total_entry + total_exit,
         station_line = fct_reorder(station_line, total),
         rank = order(total, decreasing = TRUE)
         ) %>% 
  filter(rank %in% c(1:10)) %>% 
  left_join(subway_info) %>% 
  mutate(
    station_line = fct_reorder(station_line,total)
  ) %>% 
  plot_ly(
    x = ~station_line, y = ~total, type = "bar",
    color = ~station_line, alpha = 0.5)

```
This is the top 10 buiests station during 2021-01-01 and 2021-06-30.

## Top 5 busiest station and their daily passenger over time
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
# get the top X busiest stations and get their name
top_5_str <- 
  passenger_df_time %>% 
  group_by(station_line) %>% 
  summarize(
    total_entry = sum(entry_diff_imputed,na.rm = TRUE),
    total_exit = sum(exit_diff_imputed,na.rm = TRUE),
  ) %>% 
  mutate(total = total_entry + total_exit,
         station_line = fct_reorder(station_line, total),
         rank = order(total, decreasing = TRUE)
         ) %>% 
  filter(rank %in% c(1:5)) %>% 
  pull(station_line)

# draw the line of top x lines
passenger_df_time %>% 
  filter(station_line %in% top_5_str) %>% 
  group_by(station_line,date) %>% 
  summarize(
    total_entry = sum(entry_diff_imputed,na.rm = TRUE),
    total_exit = sum(exit_diff_imputed,na.rm = TRUE)
    )%>%
  mutate(total = total_entry + total_exit) %>% 
  ggplot(aes(x = date,y= total,color = station_line))+
  geom_line() +
  geom_point() +
  transition_reveal(date)
      
```
This is the top 5 buiests station during 2021-01-01 and 2021-06-30 and their daily passenger data

# Busiest Line
### Get the busiest line
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
# if a station have more than one line, divide the passenger by the number of line
passenger_line_df <- 
  passenger_df_time %>% 
  mutate(
    entry_single = entry_diff_imputed/nchar(linename),
    exit_single = exit_diff_imputed/nchar(linename),
    total = entry_single + exit_single
    ) %>% 
  select(station_line,linename,date,total)

# function that get the passenger of specific line and date.
line_function <- function(date1,line){
 line_df <-  
  passenger_line_df %>% 
  filter(
    date == date1,
    str_detect(linename,line)
    )
 
 number = sum(pull(line_df,total),na.rm = TRUE)
 
 return(number)
}

#  a dataframe for each line in everyday 
final <- tibble(linename = str_sort(unique(unlist(strsplit(unique(passenger_df$linename),"")))))
date_series1 <- as.data.frame(date_series)
passenger_timeline_df <- 
  merge(date_series1,final)

# get the data for each line
passenger_timeline_df <- 
  passenger_timeline_df %>% 
  mutate(
    total = map2(date_series,linename,line_function),
  )

# the line rank
passenger_timeline_df %>% 
  group_by(linename) %>% 
  summarize(
    total_people = sum(unlist(total))
  ) %>% 
  mutate(
    linename = fct_reorder(linename,total_people)
  ) %>% 
  plot_ly(
    x = ~linename,y = ~total_people, color = ~linename,type = "bar" 
  )
```
This is the passenger data of each line during 2021-01-01 and 2021-06-30

### daily passenger of each line (except 1)
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
passenger_timeline_df %>%
  mutate(
    total = round(unlist(total))
  ) %>% 
  filter(linename != 1) %>% 
  ggplot(aes(x=date_series,y=total,color = linename))+
  geom_line()+
  theme_bw()+
  facet_wrap(~linename) +
  transition_reveal(date_series) +
  ease_aes('linear')
```
This is the daily passenger data of each line during 2021-01-01 and 2021-06-30(except line 1)

### daily passenger of each line (except 1)
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
passenger_timeline_df %>%
  mutate(
    total = round(unlist(total))
  ) %>% 
  filter(linename == 1) %>% 
  ggplot(aes(x=date_series,y=total,color = linename))+
  geom_line()+
  theme_bw()+
  facet_wrap(~linename) +
  transition_reveal(date_series) +
  ease_aes('linear')


```
This is the daily passenger data of line 1 during 2021-01-01 and 2021-06-30.

### Get the 10 busiest station in line 1
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
passenger_df_time %>% 
  filter(str_detect(linename, "1")) %>% 
  mutate(
    entry_single = entry_diff_imputed/nchar(linename),
    exit_single = exit_diff_imputed/nchar(linename),
  ) %>%
  group_by(station,linename) %>% 
  summarize(
    total_entry = sum(entry_single,na.rm = TRUE),
    total_exit = sum(exit_single,na.rm = TRUE),
  ) %>% 
  ungroup() %>% 
  mutate(total = total_entry + total_exit,
         station = fct_reorder(station, total),
         rank = order(total, decreasing = TRUE)) %>% 
  filter(rank %in% c(1:10)) %>% 
  ggplot(aes(x =station,y=total)) +
  geom_col()+
  coord_flip()
```
This is the top 10 busiest station in line 1 during 2021-01-01 and 2021-06-30.

##animation: line 1 station
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
animtaion_line1 <- 
  passenger_df_time %>% 
  filter(str_detect(linename, "1")) %>% 
  mutate(
    entry_single = entry_diff_imputed/nchar(linename),
    exit_single = exit_diff_imputed/nchar(linename),
    total = entry_single + exit_single
  ) %>% 
  group_by(date,station_line) %>% 
  summarize(
    total_people = sum(total)
  ) %>%
  ggplot(aes(x=station_line,y = total_people))+
  geom_col()+
  coord_flip()+
  labs(title = "{closest_state}") +
  transition_states(date)

animate(animtaion_line1, fps=1)
```
This is the daily passenger data of each station in line 1 during 2021-01-01 and 2021-06-30.

# Crime

### Crime Data
```{r echo=FALSE,tidy=TRUE,message=FALSE}
# read the crime date and adjust time variable, create unique variable for each station(station_line), count the crime data for each station in each period
subwaycrime_with_station <- 
  read_csv("data/subwaycrime_with_station.csv") %>% 
  filter(
    cmplnt_to_dt != "NA"
  ) %>% 
  mutate(
    date = mdy(cmplnt_to_dt),
      end_time = case_when(
      hms("00:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("04:00:00") ~hms("04:00:00"),
      hms("04:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("08:00:00") ~hms("08:00:00"),
      hms("08:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("12:00:00") ~hms("12:00:00"),
      hms("12:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("16:00:00") ~hms("16:00:00"),
      hms("16:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("20:00:00") ~hms("20:00:00"),
      hms("20:00:00")<=cmplnt_to_tm&cmplnt_to_tm<hms("24:00:00") ~hms("24:00:00"),
    ),
    linename = closest_line,
    station_line = str_c(closest_station,closest_line,sep = " line:"),
  ) %>% 
  select(station_line,linename,ofns_desc,pd_desc,law_cat_cd,date,end_time) %>% 
  group_by(station_line,linename,date,end_time,law_cat_cd) %>% 
  summarize(
    number = n()
  ) %>% 
  pivot_wider(
    names_from = law_cat_cd,
    values_from = number
  ) %>% 
  janitor::clean_names() %>% 
  mutate(
    felony = replace_na(felony,0),
    misdemeanor = replace_na(misdemeanor,0),
    violation = replace_na(violation,0)
  )

# set passenge data for crime data
passenger_crime <- 
  passenger_df_time %>%
  mutate(
    end_time = hours(end_time/3600)
  ) %>%
  select(station_line,linename,date,end_time,entry_diff_imputed,exit_diff_imputed)

# merge passenger and crime data
passenger_crime_df <- 
  left_join(passenger_crime,subwaycrime_with_station,by = c("station_line","linename","date","end_time")) %>% 
  mutate(
    felony = replace_na(felony,0),
    misdemeanor = replace_na(misdemeanor,0),
    violation = replace_na(violation,0),
    crime = felony + misdemeanor + violation,
    total = entry_diff_imputed+exit_diff_imputed
  )

```


### Line crime data
```{r echo=FALSE,tidy=TRUE,message=FALSE,warning=FALSE}
# function that get the passenger of specific line and date.
line_passenger_function <- function(date1,line){
 line_df <-  
  passenger_crime_df %>% 
  filter(
    date == date1,
    str_detect(linename,line)
    )
 
 number = sum(pull(line_df,total),na.rm = TRUE)
 
 return(number)
}

line_crime_function <- function(date1,line){
 line_df <-  
  passenger_crime_df %>% 
  filter(
    date == date1,
    str_detect(linename,line)
    )
 
 number = sum(pull(line_df,crime),na.rm = TRUE)
 
 return(number)
}

#  a dataframe for each line in everyday 
final <- tibble(linename = str_sort(unique(unlist(strsplit(unique(passenger_df$linename),"")))))
date_series1 <- as.data.frame(date_series)
passenger_timeline_df <- 
  merge(date_series1,final)

# get the data for each line
passenger_time_crime_line_df <- 
  passenger_timeline_df %>% 
  mutate(
    total_people = map2(date_series,linename,line_passenger_function),
    total_crime = map2(date_series,linename,line_crime_function),
  )

passenger_time_crime_line_df <- 
  passenger_time_crime_line_df%>%
  mutate(
    total_people = round(unlist(total_people)),
    total_crime = unlist(total_crime)
  ) 

# plot of passenger and crime data
time_crime <- 
  passenger_time_crime_line_df %>% 
  ggplot(aes(x = total_people,y = total_crime,color = linename,size = (total_crime/total_people))) +
  geom_point() +
  labs(title = 'Date: {frame_time}') +
  transition_time(date_series)


animate(time_crime, fps=1)
```
This is the daily passenger and crime data  of each  line  during 2021-01-01 and 2021-06-30.

```{r}
passenger_df_time %>% 
  filter(str_detect(linename, "1")) %>% 
  mutate(total = entry_diff_imputed + entry_diff_imputed) %>% 
  group_by(date) %>% 
  summarize(total_people =sum(total,na.rm=TRUE)) %>% 
  ggplot(aes(x = date, y = total_people ))+
  geom_line()
```

